Enzo+Moray: Radiation Hydrodynamics Adaptive Mesh Refinement Simulations with Adaptive Ray Tracing

### 2010-08-17 ms.tex
Adding figure for the two-source test problem.

Only in the past 15 years, computational resources have become large
enough, along with further algorithmic advances, to cope with the
requirements of three-dimensional calculations.  There are two popular
methods to solve the radiative transfer equation in three-dimensions:
%
\begin{enumerate}
\item Moment methods---The angular moments of the radiation field can
  describe its angular structure, which are related to the energy
  energy, flux, and radiation pressure \citep{Auer70, Norman98}.
  These have been implemented in conjunction with short
  characteristics \citep{Stone92_RT}, with long characteristics
  \citep{Finlator08}, and with a variable Eddington tensor in the
  optically thin limit \citep[OTVET;][]{Gnedin01_OTVET, Petkova09}.
  Moment methods have the advantage of being fast, being independent
  of the number of radiation point sources.  However, they are
  diffusive and result in incorrect shadows in some situations.
\item Ray tracing---Radiation can be propagated along rays that extend
  through the compuational grid \citep[e.g.][]{Mellema98, Razoumov99,
    Abel99, Ciardi00, Ritzerveld03, Whalen05, Rijkhorst05, Mellema06,
    Alvarez06, Trac07, Krumholz07_ART, Paardekooper10} or particle set
  \citep[e.g.][]{Susa06, Johnson07, Pawlik08, Altay08, Hasegawa09,
    Pawlik10}.  In general, these methods are very accurate but
  computationally expensive because the radiation field must be well
  sampled by the rays with respect to the spatial resolution of the
  grid or particles.
\end{enumerate}
%
Until the mid-2000's the vast majority of the three-dimensional
calculations were performed with static density distributions.  One
example is calculating cosmological reionization by post-processing of
density fields from N-body simulations \citep{Ciardi01, Sokasian01,
  McQuinn05, Iliev07}.  Any hydrodynamical response to the radiation
field was thus ignored.  Several radiative transfer codes were
compared in four purely radiative transfer tests in \citet[hereafter
RT06]{RT06}.  Only recently has the radiative transfer equation been
coupled to the hydrodynamics in three-dimensions.  In the second
comparision paper \citep[hereafter RT09]{Iliev09}, results from these
radiation hydrodynamics codes were compared.  Even more rare are ones
that couple it with magneto-hydrodynamics
\citep[e.g.][]{Krumholz07_ART}.  The tests in RT06 and RT09 were kept
relatively simple to ease the comparison.

@@

A supernova blastwave being irradiated by a nearby star is a likely
occurence in massive-star forming regions.  In this test, we set up an
idealized test that mimics this scenario.  The problem is similar to
Test 5 with two changes --- (1) the domain has been decreased to
$L_{\rm box} = 1.5$ kpc; and (2) a mild blastwave is initialized in
the center of the domain.  We use 3 levels of AMR with a base grid of
64$^3$, which is refined if the density slope is greater than 0.4.
The blastwave is initially in the free expansion phase with a total
energy of $10^{50}$ erg and total mass of $100 \Ms$, corresponding to
$E = 315$ eV per particle or $E/k_b = 3.66 \times 10^6$ K.  The
blastwave is setup with a linearly increasing velocity profile
according to \citet{Truelove99} with $v_{\rm max} = 130$ km s$^{-1}$
and a radius of $L_{\rm box}/50 = 30$ pc.  The blastwave initially has
a uniform density of $\rho_0 = 0.06 \cubecm$.  The initial kinetic
energy is converted into thermal energy as the shock propagates away,
and the blastwave transitions into the Sedov-Taylor phase at $r \sim
100$ pc and $t \sim 225$ kyr after it has accumulated an equal amount
of ambient material into the shock.  The simulation is run for 25 Myr.

@@

@@

The \ion{H}{2} regions grow to $r = 4$ kpc where they overlap.  Then
the two sources are enveloped in a common, elongated \ion{H}{2}
region.

>>>>

Only in the past 15 years, computational resources have become large
enough, along with further algorithmic advances, to cope with the
requirements of three-dimensional calculations.  There are two popular
methods to solve the radiative transfer equation in three-dimensions:
%
\begin{enumerate}
\item Moment methods---The angular moments of the radiation field can
  describe its angular structure, which are related to the energy
  energy, flux, and radiation pressure \citep{Auer70, Norman98}.
  These have been implemented in conjunction with short
  characteristics \citep{Stone92_RT}, with long characteristics
  \citep{Finlator08}, with a variable Eddington tensor in the
  optically thin limit \citep[OTVET;][]{Gnedin01_OTVET, Petkova09},
  and with an M1 closer relation \citep{Gonzalez07, Aubert08}.  Moment
  methods have the advantage of being fast, being independent of the
  number of radiation point sources.  However, they are diffusive and
  result in incorrect shadows in some situations.
\item Ray tracing---Radiation can be propagated along rays that extend
  through the compuational grid \citep[e.g.][]{Mellema98, Razoumov99,
    Abel99, Ciardi00, Ritzerveld03, Whalen05, Rijkhorst05, Mellema06,
    Alvarez06, Trac07, Krumholz07_ART, Paardekooper10} or particle set
  \citep[e.g.][]{Susa06, Johnson07, Pawlik08, Altay08, Hasegawa09,
    Pawlik10}.  In general, these methods are very accurate but
  computationally expensive because the radiation field must be well
  sampled by the rays with respect to the spatial resolution of the
  grid or particles.
\end{enumerate}
%
Until the mid-2000's the vast majority of the three-dimensional
calculations were performed with static density distributions.  One
example is calculating cosmological reionization by post-processing of
density fields from N-body simulations \citep{Ciardi01, Sokasian01,
  McQuinn05, Iliev07}.  Any hydrodynamical response to the radiation
field was thus ignored.  Several radiative transfer codes were
compared in four purely radiative transfer tests in \citet[hereafter
RT06]{RT06}.  Only recently has the radiative transfer equation been
coupled to the hydrodynamics in three-dimensions.  In the second
comparision paper \citep[hereafter RT09]{Iliev09}, results from these
radiation hydrodynamics codes were compared.  Even more rare are ones
that couple it with magneto-hydrodynamics
\citep[e.g.][]{Krumholz07_ART}.  The tests in RT06 and RT09 were kept
relatively simple to ease the comparison.

@@

A supernova blastwave being irradiated by a nearby star is a likely
occurence in massive-star forming regions.  In this test, we set up an
idealized test that mimics this scenario.  The problem is similar to
Test 5 with two changes --- (1) the domain has been decreased to
$L_{\rm box} = 1.5$ kpc; and (2) a blastwave is initialized in the
center of the domain.  We use 3 levels of AMR with a base grid of
64$^3$, which is refined if the density slope is greater than 0.4.
The blastwave is initially in the free expansion phase with a total
energy of $10^{50}$ erg and total mass of $100 \Ms$, corresponding to
$E = 315$ eV per particle or $E/k_b = 3.66 \times 10^6$ K.  The
blastwave is setup with a linearly increasing velocity profile
according to \citet{Truelove99} with $v_{\rm max} = 130$ km s$^{-1}$
and a radius of $L_{\rm box}/50 = 30$ pc.  The blastwave initially has
a uniform density of $\rho_0 = 0.06 \cubecm$.  The initial kinetic
energy is converted into thermal energy as the shock propagates away,
and the blastwave transitions into the Sedov-Taylor phase at $r \sim
100$ pc and $t \sim 225$ kyr after it has accumulated an equal amount
of ambient material into the shock.  The simulation is run for 25 Myr.

@@

\begin{figure}[t]
  \epsscale{0.6}
  \plotone{fig/two_sources.eps}
  \caption{\label{fig:twosrc} Slice of neutral fraction at $t = 500$
    Myr through the sources in the consolidated \ion{H}{2} test.
    There are no artifacts associated with rays being emitted from two
    sources.  Both of the ionized regions are spherically symmetric
    before they overlap.}
  \epsscale{1}
\end{figure}

@@

The \ion{H}{2} regions grow to $r = 4$ kpc where they overlap.  Then
the two sources are enveloped in a common, elongated \ion{H}{2}
region.  To illustrate this, we show the neutral fraction in Figure
\ref{fig:twosrc}.  Our method keeps spherical symmetry close to the
individual sources, and there are no perceptible artifacts from having
multiple sources.
